Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-05 manuscript-chicago.Rmd
Removing language for word count

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
appendix          :

@@

Today, social media replicates and amplifies this phenomenon during moments of
social unrest. Amid the cacophony of messages during political upheaval---such
as the Arab Spring---consistent themes emerge through hashtags, retweets, and
the repeated mentions of concepts such as regime, security, demonstration, and
human rights. The cohesiveness of these themes (i.e., whether movement
participants are metaphorically singing in one voice) reflects the real-world
events happening contemporaneously with collective action in social movements or
civil conflicts.

@@

In December 2010, a Tunisian man named Mohamed Bouazizi was harassed and
unjustifiably stripped of his vendor cart by municipal officials. His
humiliation led him to set himself on fire in protest to the authoritarian rule,
ultimately leading to his death several days later. His actions served as a
catalyst for the Tunisian Revolution, and one week after his death, Zine El
Abidine Ben Ali---the longstanding authoritarian president---fled the country.
The events of Tunisia rapidly spread to surrounding countries also subject to
unjust authoritarian rule, including Yemen, Egypt, Libya, and Syria.

@@

Massive online social cohesion can have real-world social manifestations. Social
media platforms provide one way to enable this kind of online social cohesion by
offering distributed information-gathering and real-time information
dissemination. Given its accessibility and proficiency as an information
exchange tool, Twitter has provided individuals involved in social movements an
opportunity to complement or even replace real-world social mobilization. For
example, during the 2012 Italian protests for global economic change, the riots
in Rome were mirrored by Twitter activity both on the day of the protests and in
the days following them [@vicari2013public]. Tweeting was determined to be more
effective locally in discussing real-world, real-time events than traditional
mainstream media platforms. As more and more individuals rely on social media to
obtain local news and information, it becomes increasingly important to
understand just how these micro-blogs reflect real-world action.

During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms.

@@

Social media generates critical information about events, including whether and
how a demonstration will take place, how many people participate, and how those
in power (e.g., state, military, security apparatus) respond. Similarly, these
media platforms alert us to more global shifts within a conflict, such as when
fighting moves to a new city or region, the number of casualties and wounded,
and the types of weaponry used by parties to the conflict [@zeitzoff2017social].
Social media is not entirely event-driven, however. In monitoring social media
channels, government forces may respond to the perception that citizens are
gaining strength and sympathy. Governments have been shown to be quite tolerant
of dissent so long as it does not metamorphose into collective action
[@king2013censorship].

@@

\setlength{\leftskip}{0pt}
\end{singlespace}
\noindent With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes [-@gohdes2015pulling] notes that regimes can attempt to thwart the growth
of social protest online and in the streets by shutting down the Internet.
During these times, however, she finds that military repression increases.
However, government information about the citizens decreases; as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

@@

Previous research on social movements supports the idea that participants are
more likely to join movements when they know and have a personal connection to
someone else already involved [@mcadam1986recruitment]. The type of collective
contentious behavior that manifests in society is a function of the available
political opportunity structures. The political opportunity structure are
“comprised of *(sic)* specific configurations of resources, institutional
arrangements and historical precedents for social mobilization, which facilitate
the development of protest movements in some instances and constrain them in
others” [@kitschelt1986political]. What emerges as strategies and tactics in
contentious politics are, in essence, a function of what the system makes
possible. This includes the number of people and their disposition (i.e.
nonviolent or violent), as well as their internal structure. The more cohesive a
group is, the greater threat they pose to the establishment.

@@

```{r cohesion-in-tweets, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "A visual representation of cohesion in tweets. Content word overlap indicates a shared relationship among tweets. The shared words--either identical or semantically similar--constitute a theme around which cohesion is occuring.", out.width='100%'}

@@

An additional focus of the current work is to introduce cross-recurrence
quantification analysis---a longstanding method from physics that has since
become influential in a variety of other fields---to political science. We do so
by providing detailed descriptions of the methods and providing clear directions
about interpretation and significance testing even with relatively small-*n* or
case studies. After demonstrating these methods' utility with this dataset, we
close the paper with specific suggestions to political scientists interested in
incorporating these methods into their work.

>>>>

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
appendix          :

@@

Today, social media replicates and amplifies this phenomenon during moments of
social unrest. Amid the cacophony of messages during political upheaval---such
as the Arab Spring---consistent themes emerge through hashtags, retweets, and
the repeated mentions of concepts such as regime, demonstration, and
human rights. The cohesiveness of these themes (i.e., whether movement
participants are metaphorically singing in one voice) reflects the real-world
events happening contemporaneously with collective action in social movements or
civil conflicts.

@@

In December 2010, a Tunisian man named Mohamed Bouazizi was unjustifiably
stripped of his vendor cart by municipal officials. His humiliation led him to
set himself on fire in protest to the authoritarian rule, leading to his death
several days later. His actions served as a catalyst for the Tunisian
Revolution, and one week after his death, Zine El Abidine Ben Ali---the
longstanding authoritarian president---fled the country. The events of Tunisia
rapidly spread to surrounding countries also subject to unjust authoritarian
rule, including Egypt, Libya, and Syria.

@@

Massive online social cohesion can have real-world social manifestations. Social
media platforms provide one way to enable this kind of online social cohesion by
offering distributed information-gathering and real-time information
dissemination. Given its accessibility and proficiency as an information
exchange tool, Twitter has provided individuals involved in social movements an
opportunity to complement or even replace real-world social mobilization. For
example, during the 2012 Italian protests for global economic change, the riots
in Rome were mirrored by Twitter activity both on the day of the protests and in
the days following them [@vicari2013public]. Tweeting was determined to be more
effective locally in discussing real-world, real-time events than traditional
mainstream media platforms. As more and more individuals rely on social media to
obtain local news and information, it becomes increasingly important to
understand just how these micro-blogs reflect real-world action. Some argue that
these mass movements would not have been possible without efficient large-scale
coordination, both physically on the ground and virtually on social media
platforms.

@@

Social media generates critical information about events, including whether and
how a demonstration will take place, how many people participate, and how those
in power (e.g., state, military, security apparatus) respond. Similarly, these
media platforms alert us to more global shifts within a conflict, such as when
fighting moves to a new city or region, the number of casualties and wounded,
and the types of weaponry used by parties [@zeitzoff2017social]. In monitoring
social media channels, government forces may respond to the perception that
citizens are gaining strength and sympathy. Governments have been shown to be
quite tolerant of dissent so long as it does not metamorphose into collective
action [@king2013censorship].

@@

\setlength{\leftskip}{0pt}
\end{singlespace}
\noindent With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes [-@gohdes2015pulling] notes that regimes can attempt to thwart the growth
of social protest online and in the streets by shutting down the Internet.
During these times, however, they find that military repression increases.
However, government information about the citizens decreases; as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

@@

The type of collective contentious behavior that manifests in society is a
function of the available political opportunity structures. The political
opportunity structure are “comprised of *(sic)* specific configurations of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others” [@kitschelt1986political]. What emerges
as strategies and tactics in contentious politics are, in essence, a function of
what the system makes possible. This includes the number of people and their
disposition (i.e. nonviolent or violent), as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.

@@

```{r cohesion-in-tweets, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "A visual representation of cohesion in tweets. The shared content words--either identical or semantically similar--constitute a theme around which cohesion is occuring.", out.width='100%'}

@@

An additional focus of the current work is to introduce cross-recurrence
quantification analysis---a longstanding method from physics that has since
become influential in a variety of other fields---to political science. We do so
by providing detailed descriptions of the methods and providing clear directions
about interpretation and significance testing even with relatively small-*n* or
case studies. After demonstrating these methods' utility with this dataset, we
close the paper with specific suggestions to political scientists interested in
these methods.
